---
title: "Final_Report"
format: html
---

# 1. Introduction

-   what is this project about

-   what are our research questions maybe

# 2. Data Tidying

```{r setup, include=FALSE}

#DO NOT REMOVE, this is to set the global message settings and library imports for every single code block afterwards.

library(tidyverse)
library(readxl)

knitr::opts_chunk$set(message = FALSE)
```

## 2.1 Principle 1: Each type of case must have its own tibble. 

**Principle 1 states that each identified type of case be stored in its own tibble.** A tibble representing a specific case type should function as a clear unit of analysis and must not contain duplicated or redundant information. This often involves using functions such as `select()` to keep only relevant variables and `distinct()` to ensure that each observational unit appears only once. Storing data in such way improves storage efficiency, enhances consistency, and minimizes the risk of errors during table joins.

To ensure principle 1, there are **two key steps**:

1.   [Identify the different types of cases.]{.underline}\
    To do this, we first need to identify the primary key in each data set, because cases are defined as combinations of the primary key.

2.   [Choose the variables we need and split the data into separate tibbles.]{.underline}\
    Each tibble should include only the variables that describe that specific type of case.

### 2.1.1 Monthly Survey

**2.1.1.1 What do monthly survey questionnaires contain?**

The monthly survey is a compilation of survey responses collected across different months. Monthly survey data sets are stored in different CSV files per month (e.g. `SurveyMonthly.2010_04.csv`). Each data set has a similar structure, with first few rows consist of participants' metadata, followed by survey responses that vary month by month.

There is a file called `SurveyMonthlyQuestionKey.csv`, which listed out all the scale names, scale items, and in which monthly survey were they being asked.However, based on our research questions, some of the monthly data sets are irrelevant. We will only take the data sets that are relevant to our research questions into further data tidying process.

First, we imported all the monthly survey data sets into separate data frames:

```{r}
#creating tibbles for monthly survey per month

monthly_files <- c(
  "InitialSurvey.csv",
  "SurveyMonthly.2010_04.csv",
  "SurveyMonthly.2010_05.csv",
  "SurveyMonthly.2010_07.csv",
  "SurveyMonthly.2010_10.csv",
  "SurveyMonthly.2010_11.csv",
  "SurveyMonthly.2010_12.csv",
  "SurveyMonthly.2011_03.csv",
  "SurveyMonthly.2011_04.csv"
)

monthly_names <- c(
  "initial_df",
  "month1004_df",
  "month1005_df",
  "month1007_df",
  "month1010_df",
  "month1011_df",
  "month1012_df",
  "month1103_df",
  "month1104_df"
)

for(i in 1:length(monthly_files)) {
  assign(monthly_names[i], read.csv(monthly_files[i]))
}
```

**2.1.1.2 Primary key and cases**

As mentioned above, apart from survey responses, all monthly survey data sets included participants' metadata, such as `participantID`, `start date` and `end date`. Participants' demographics were not collected every month, as it is listed in the `SurveyMonthlyQuestionKey.csv`. We deem that `participantID` is the primary key for every data set, beneath is one example of how we check the primary key across all monthly data sets.

```{r}
#check whether participantID is the primary key for monthly survey in April, 2010

month1004_df |> 
  count(participantID) |> 
  filter(n>1)
```

We used the same code across different monthly data sets and confirmed that `participantID` is indeed the primary key for all of them. Thus, `participantID` will be included in all separate tibbles we created to act as the foreign key.

However, between different months, the survey questions do overlap according to `SurveyMonthlyQuestionKey.csv`. For example, for the research objectives "friendship network change", the relevant scale is called "group involvement", which was measured by several items across different months. For items like "Formal group involvement", it was only measured once in the initial survey, while for items like "Formal groups participation changed“ and "Formal groups new involvement“, it is an update to previous survey responses, thus were measured in both April and May, 2010, as it is listed below.

```{r}
#extract relevant information for "group involvement scale"

survey_key = read_excel("SurveyMonthlyQuestionKey.xlsx")
group_itemskey = survey_key |> 
  slice(c(168:171, 186:195))
view(group_itemskey)
```

In the next step, we extract items like "Formal groups participation changed“ and "Formal groups new involvement“ from different monthly survey and store them in different tibbles (e.g. `mar_group`, `apr_group`, as mentioned in the following paragraphs). Therefore, the cases we are using contains both (participant) and (participant, month), with the former one focused on participants' demographics, and the latter one for variables that are related to our research questions.

**2.1.1.3 Creating tibbles for different cases**

***Initial Survey/March, 2010***

This data sets contains both participants' metadata, demographics and responses to survey items. Cases for this data set are (participant) and (participant, month).

**a. Demographics**

For the case (participant), we create separate tibble for demographics from initial survey. More specifically, the demographic tibble only includes participants' ID, and participants' gender. Due to privacy issues, sensitive information such as apartment number, name, cell phone number, etc. were already removed or anonymised in the original data sets, thus were removed from the tibble. Information like start date and end date are not characteristics of the participants, thus were also removed from the tibble.

```{r}
mar_demo = initial_df |> 
  select(c(1,14)) |> 
  distinct() |> 
  arrange(participantID)
```

**b. Group participation**

```{r}
mar_group = initial_df |> 
  select(c(1, 53:76)) |> 
  distinct() |> 
  arrange(participantID)
```

**c. Communication with friends**

```{r}
mar10_commu = initial_df |> 
  select(c(1, 277:280)) |> 
  distinct() |> 
  arrange(participantID)
```

***Monthly survey April, 2010***

**a. Big 5**

```{r}
apr_big5 = month1004_df |> 
  select(c(1, 11:54)) |> 
  distinct() |> 
  arrange(participantID)
```

**b. Group participation**

```{r}
apr_group = month1004_df |> 
  select(c(1, 55:114)) |> 
  distinct() |> 
  arrange(participantID)
```

***Monthly survey May, 2010***

**a. Group participation**

```{r}
may_group = month1005_df |>
  select(c(1, 73:132)) |> 
  distinct() |> 
  arrange(participantID)
```

**b. Well-being**

```{r}
may_well_being = month1005_df |> 
  select(c(1, 32:72)) |> 
  distinct() |> 
  arrange(participantID)
```

***Monthly survey October, 2010***

**a. Aerobic exercise actual**

```{r}
oct_exe_act = month1010_df |>    
  select(c(1, 62)) |>    
  distinct() |>    
  arrange(participantID)
```

**b. Sleep actual**

```{r}
oct_sleep_act = month1010_df |> 
  select(c(1, 80)) |> 
  distinct() |> 
  arrange(participantID)
```

***Monthly survey March, 2011***

**Communication with friends**

```{r}
mar11_commu = month1103_df |> 
  select(c(1, 26:30)) |> 
  distinct() |> 
  arrange(participantID)
```

### 2.1.2 Location

Location every 30 minutes (`Location.csv.bz2`), with each record including the participant ID associated with the data-collecting mobile phone (`participantID`), the date-time when the location is recorded (`date`), the the 68% confidence offset of the estimation from true location(`accuracy`, in meters), and the estimated location affine-transformed (`x`, `y`).

```{r}
#create a tibble for Location.csv, and have a preview of the data set

location_df = read.csv("Location.csv") |> arrange(participantID)
head(location_df)
```

We tried to define the primary key, however, even if we combine all the column names together, the combination itself still cannot identify every single row uniquely, as it is shown in the following code:

```{r}
#check whether (participantID, date, accuracy, x, y) is the primary key

location_df |> 
  count(participantID, date, accuracy, x, y) |> 
  filter(n>1) |> 
  head()
```

We deemed that the duplicates were produced due to faulty data collection methods after reviewing the data description. Thus, we decided to remove duplicated data that has the exact value under every variable:

```{r}
location_df = location_df |> distinct()
```

In the new data frame, we found that (`participantID`, `date`) is the primary key, as it is shown in the following code:

```{r}
#check whether (participantID, date) is the primary key

location_df |> 
  count(participantID, date) |> 
  filter(n>1)
```

In the final data frame, for the case participant ID, no characteristic was involved. For the case (`participantID`, `date`), we created a separate tibble according to principle 1:

```{r}
date_coordinates = location_df |> 
  select(participantID, date, x, y) |> 
  distinct() |> 
  arrange(participantID)
```



## 2.2 Principle 2: Each observation must have its own row. (Draft ver.)

what is principle 2... For principle 2, the researcher needed to ensure that no multiple observations were stacked into the same row. In order to achieve that, the researcher first examined all 11 tibbles created from the first step to spot repeated information. Then, the researcher should delete repeated information among different rows.

### 2.2.1 Monthly Survey

**a. Demographics**

```{r}
#check to count the distinct values within Gender

mar_demo |> 
  count(Gender)
```

```{r}
# recode the values in Gender to be consistent and use the same expressions

mar_demo <- mar_demo |> 
  transmute(participantID = participantID,
            Gender =
              recode(Gender,
                     "F" = "Female",
                     "f" = "Female",
                     "female" = "Female",
                     "M" = "Male",
                     "m" = "Male",
                     "male" = "Male"))

#check to count the distinct values within Gender

mar_demo |> 
  count(Gender)
View(mar_demo)
```

There were in total 86 + 56 = 142 participants, but only 141 rows were found in mar_demo. The researcher then explored the missing participant and filled in their data with NA value.

```{r}
# check if there are participants missing from the demographics df.

all_participants <- c(paste0("fa10-01-", sprintf("%02d", 1:86)),
                      paste0("sp10-01-", sprintf("%02d", 1:56)))
setdiff(all_participants, mar_demo$participantID)

# add the row for the missing participant fa10-01-84.

mar_demo <- mar_demo |> 
  add_row(Gender = NA, participantID = 'fa10-01-84') |> 
  arrange(participantID)
```

**b. Group Participation**

```{r}
# delete rows with all NA values because they provide essentially no information.

mar_group <- mar_group |> 
  filter(!if_all(-participantID, ~is.na(.)))
```

There should be 56 participants with answers for Group Participation, but only 55 rows of data were there. The researcher then explored the missing participant.

```{r}
# explore which participant is missing from the data frame.

all_participants <- c(paste0("sp10-01-", sprintf("%02d", 1:56)))
setdiff(all_participants, mar_group$participantID)
```

In addition, the researcher noticed that the NA values were not consistently represented in all cells, so the researcher decided to recode values to retain consistency.

```{r}
# fill all empty cells with NA values

mar_group <- mar_group |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  distinct()
```

**c. Communication with friends**

```{r}
# delete rows with all NA values because they provide essentially no information.

mar10_commu <- mar10_commu |> 
  filter(!if_all(-participantID, ~is.na(.)))

# fill all empty cells with NA values

mar10_commu <- mar10_commu |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  distinct()
```

```{r}
# explore how many participants are missing from the mar10_commu df.

all_participants <- c(
  paste0("sp10-01-", sprintf("%02d", 1:56)),
  paste0("fa10-01-", sprintf("%02d", 1:40))
)
setdiff(all_participants, mar10_commu$participantID) |> 
  length()
```

***Monthly survey April, 2010***

**a. Big 5**

```{r}
# fill all empty cells with NA values

apr_big5 <- apr_big5 |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  distinct()
```

```{r}
# explore how many participants are missing from the apr_big5 df.

all_participants <- c(
  paste0("sp10-01-", sprintf("%02d", 1:56))
)
setdiff(all_participants, apr_big5$participantID) |> 
  length()
View(apr_big5)
```

**b. Group participation**

```{r}
# fill all empty cells with NA values

apr_group <- apr_group |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  distinct()

# delete rows with all NA values because they provide essentially no information.

apr_group <- apr_group |> 
  filter(!if_all(-participantID, ~is.na(.)))
```

```{r}
# explore how many participants are missing from the apr_group df.

all_participants <- c(
  paste0("sp10-01-", sprintf("%02d", 1:56))
)
setdiff(all_participants, apr_group$participantID) |> 
  length()
```

```{r}
# explore how many participants are missing from the apr_group df.

all_participants <- c(
  paste0("sp10-01-", sprintf("%02d", 1:56))
)
setdiff(all_participants, apr_group$participantID) |> 
  length()
```

The researcher spotted several repeated values in the columns and decided to pivot the columns into multiple rows for each different observation.

```{r}
# change the data type into character

apr_group <- apr_group |>
  mutate(across(-participantID, as.character))

# change the data type for group number into integar

apr_group <- apr_group |>
  mutate(group_number = as.integer(group_number))
```

***Monthly survey May, 2010***

**a. Group participation**

```{r}

may_group = month1005_df |>
  select(c(1, 73:132)) |> 
  distinct() |> 
  arrange(participantID)
```

```{r}
# fill all empty cells with NA values

may_group <- may_group |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  distinct()

# delete rows with all NA values because they provide essentially no information.

may_group <- may_group |> 
  filter(!if_all(-participantID, ~is.na(.)))

# explore how many participants are missing from the apr_group df.

all_participants <- c(
  paste0("sp10-01-", sprintf("%02d", 1:56))
)
setdiff(all_participants, may_group$participantID) |> 
  length()
```

**b. Well-being**

```{r}
# fill all empty cells with NA values

may_well_being <- may_well_being |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  distinct()

# delete rows with all NA values because they provide essentially no information.

may_well_being <- may_well_being |> 
  filter(!if_all(-participantID, ~is.na(.)))

# explore how many participants are missing from the may_well_being df.

all_participants <- c(
  paste0("sp10-01-", sprintf("%02d", 1:56))
)
setdiff(all_participants, may_well_being$participantID) |> 
  length()
```

***Monthly survey October, 2010***

**a. Aerobic exercise actual**

```{r}
# fill all empty cells with NA values

oct_exe_act <- oct_exe_act |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  distinct()

# delete rows with all NA values because they provide essentially no information.

oct_exe_act <- oct_exe_act |> 
  filter(!if_all(-participantID, ~is.na(.)))

# explore how many participants are missing from the oct_exe_act df.

all_participants <- c(
  paste0("sp10-01-", sprintf("%02d", 1:56)),
  paste0("fa10-01-", sprintf("%02d", 1:40))
)
setdiff(all_participants, oct_exe_act$participantID) |> 
  length()
```

**b. Sleep actual**

```{r}
# fill all empty cells with NA values

oct_sleep_act <- oct_sleep_act |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  distinct()

# delete rows with all NA values because they provide essentially no information.

oct_sleep_act <- oct_sleep_act |> 
  filter(!if_all(-participantID, ~is.na(.)))

# explore how many participants are missing from the oct_sleep_act df.

all_participants <- c(
  paste0("sp10-01-", sprintf("%02d", 1:56)),
  paste0("fa10-01-", sprintf("%02d", 1:40))
)
setdiff(all_participants, oct_sleep_act$participantID) |> 
  length()
```

***Monthly survey March, 2011***

**Communication with friends**

```{r}

# delete rows with all NA values because they provide essentially no information.

mar11_commu <- mar11_commu |> 
  filter(!if_all(-participantID, ~is.na(.)))

# fill all empty cells with NA values

mar11_commu <- mar11_commu |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  distinct()

# explore how many participants are missing from the mar11_commu df.

all_participants <- c(
  paste0("sp10-01-", sprintf("%02d", 1:56)),
  paste0("fa10-01-", sprintf("%02d", 1:40))
)
setdiff(all_participants, mar11_commu$participantID) |> 
  length()
```

### 2.1.2 Location

```{r}
# delete possible repeated rows

date_coordinates <- date_coordinates |> 
  distinct(participantID, date, x, y, .keep_all = TRUE)

# fill all empty cells with NA values

date_coordinates <- date_coordinates |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  distinct()

# delete rows with all NA values because they provide essentially no information.

date_coordinates <- date_coordinates |> 
  filter(!if_all(-participantID, ~is.na(.)))

# explore how many participants are missing from the date_coordinates df.

all_participants <- c(
  paste0("sp10-01-", sprintf("%02d", 1:56)),
  paste0("fa10-01-", sprintf("%02d", 1:40))
)
setdiff(all_participants, date_coordinates$participantID) |> 
  length()
```


