---
title: "Final_Report"
format:
  html:
    toc: true          
    toc-location: left 
---

# 1.Introduction

The Friends and Family study was developed to better understand how people make decisions in their everyday lives, with a particular interest in how social relationships shape those choices and whether certain personal or social tools might help people make more effective decisions. The study took place in a residential community for young families located next to a major research university in North America. All households consisted of couples, and at least one partner in each couple was affiliated with the university. More than 130 adults participated during 2010 across two recruitment phases, forming a relatively small, closely connected community where many families also had children.

To capture detailed information about daily behaviour and interactions, participants used their personal Android smartphones that automatically recorded call logs, GPS location data, and SMS activity. These behavioural traces were combined with repeated self-report measures to gain insight into personal characteristics, attitudes, routines, and well-being. Participants completed a baseline personality survey, followed by a series of monthly and weekly questionnaires, as well as daily reports related to activities, lifestyle habits, and mood.

## 1.1 Research Questions

Each group member selected an individual research question to explore using the tidied data set. The research questions that group members will further study are the following:

1.  How are sleep patterns and aerobic exercise levels related, and does this relationship differ by gender?

2.  How does a friendship network change over time, and how might those changes relate to individual characteristics?

3.  How does the frequency of different network types—such as friendship, fitness, and call/SMS interactions—relate to participant characteristics?

4.  How do movement patterns vary between individuals who move frequently and those who move infrequently, and how might these patterns be connected to well-being?

5.  How does communication with remote friends shift over time, and how do these communication patterns relate to personal attributes?



# 2. Data Tidying

```{r setup, include=FALSE}

#DO NOT REMOVE, this is to set the global message settings and library imports for every single code block afterwards.

#| echo: false
#| warning: false

library(tidyverse)
library(readxl)

knitr::opts_chunk$set(message = FALSE)
```

## 2.1 Principle 1: Each type of case must have its own tibble.

**Principle 1 states that each identified type of case be stored in its own tibble.** A tibble representing a specific case type should function as a clear unit of analysis and must not contain duplicated or redundant information. This often involves using functions such as `select()` to keep only relevant variables and `distinct()` to ensure that each observational unit appears only once. Storing data in such way improves storage efficiency, enhances consistency, and minimizes the risk of errors during table joins.

To ensure principle 1, there are **two key steps**:

1.  [Identify the different types of cases.]{.underline}\
    To do this, we first need to identify the primary key in each data set, because cases are defined as combinations of the primary key.

2.  [Choose the variables we need and split the data into separate tibbles.]{.underline}\
    Each tibble should include only the variables that describe that specific type of case.

### 2.1.1 Monthly Survey

#### **2.1.1.1 What do monthly survey questionnaires contain?**

The monthly survey is a compilation of survey responses collected across different months. Monthly survey data sets are stored in different CSV files per month (e.g. `SurveyMonthly.2010_04.csv`). Each data set has a similar structure, with first few rows consist of participants' metadata, followed by survey responses that vary month by month.

There is a file called `SurveyMonthlyQuestionKey.csv`, which listed out all the scale names, scale items, and in which monthly survey were they being asked.However, based on our research questions, some of the monthly data sets are irrelevant. We will only take the data sets that are relevant to our research questions into further data tidying process.

First, we imported all the monthly survey data sets into separate data frames:

```{r}
#creating tibbles for monthly survey per month

monthly_files <- c(
  "InitialSurvey.csv",
  "SurveyMonthly.2010_04.csv",
  "SurveyMonthly.2010_05.csv",
  "SurveyMonthly.2010_07.csv",
  "SurveyMonthly.2010_10.csv",
  "SurveyMonthly.2010_11.csv",
  "SurveyMonthly.2010_12.csv",
  "SurveyMonthly.2011_03.csv",
  "SurveyMonthly.2011_04.csv"
)

monthly_names <- c(
  "initial_df",
  "month1004_df",
  "month1005_df",
  "month1007_df",
  "month1010_df",
  "month1011_df",
  "month1012_df",
  "month1103_df",
  "month1104_df"
)

for(i in 1:length(monthly_files)) {
  assign(monthly_names[i], read.csv(monthly_files[i]))
}
```

#### **2.1.1.2 Primary key and cases**

As mentioned above, apart from survey responses, all monthly survey data sets included participants' metadata, such as `participantID`, `start date` and `end date`. Participants' demographics were not collected every month, as it is listed in the `SurveyMonthlyQuestionKey.csv`. We deem that `participantID` is the primary key for every data set, beneath is one example of how we check the primary key across all monthly data sets.

```{r}
#check whether participantID is the primary key for monthly survey in April, 2010

month1004_df |> 
  count(participantID) |> 
  filter(n>1)
```

We used the same code across different monthly data sets and confirmed that `participantID` is indeed the primary key for all of them. Thus, `participantID` will be included in all separate tibbles we created.

However, between different months, the survey questions do overlap according to `SurveyMonthlyQuestionKey.csv`. For example, for the research objectives "friendship network change", the relevant scale is called "group involvement", which was measured by several items across different months. For items like "Formal group involvement", it was only measured once in the initial survey, while for items like "Formal groups participation changed“ and "Formal groups new involvement“, it is an update to previous survey responses, thus were measured in both April and May, 2010, as it is listed below.

```{r}
#extract relevant information for "group involvement scale"

survey_key = read_excel("SurveyMonthlyQuestionKey.xlsx")
group_itemskey = survey_key |> 
  slice(c(168:171, 186:195))

#present the information in a table

knitr::kable(head(group_itemskey), format = "pipe") 
```

In the next step, we extract items like "Formal groups participation changed“ and "Formal groups new involvement“ from different monthly survey and store them in different tibbles (e.g. `mar_group`, `apr_group`, as mentioned in the following paragraphs). Although cases we can identify primarily just (participant) itself, however, when it comes to survey responses, they were not characteristics for participants. Instead, survey responses are characteristics of (what) participants (answered per month). Therefore, it only makes sense if we include both (participant) and (participant, month) as our cases, with the former one focused on participants' demographics, and the latter one for variables that are related to our research questions.

#### **2.1.1.3 Creating tibbles for different cases**

In the following parts, according to `SurveyMonthlyQuestionKey.csv`, we extract survey items that measure our research variables into separate tibbles.

##### **Initial Survey/March, 2010**

This data sets contains both participants' metadata, demographics and responses to survey items. Cases for this data set are (participant) and (participant, month).

*Demographics*

For the case (participant), we create separate tibble for demographics from initial survey. More specifically, the demographic tibble only includes participants' ID, and participants' gender. Due to privacy issues, sensitive information such as apartment number, name, cell phone number, etc. were already removed or anonymised in the original data sets, thus were removed from the tibble. Information like start date and end date are not characteristics of the participants, thus were also removed from the tibble.

```{r}
mar_demo = initial_df |> 
  select(c(1,14)) |> 
  distinct() |> 
  arrange(participantID)

#present an overview of the tibble

knitr::kable(head(mar_demo), format = "pipe") 
```

*Group participation*

Group participantion was consisted of two parts, in the initial survey, only the formal/informal group participation was recorded. In the follow up surveys, items like whether participants stopped participating groups they mentioned in the initial survey, were also recorded. This variable was recorded across March, April and May (differently).

```{r}
mar_group = initial_df |> 
  select(c(1, 53:76)) |> 
  distinct() |> 
  arrange(participantID)

#present an overview of the tibble

knitr::kable(head(mar_group), format = "pipe") 
```

*Communication with friends*

Communication with friends was defined as the frequency of communicating with friends through different remote channels. This scale is measured repetitively across March, 2010 and March 2011.

```{r}
mar10_commu = initial_df |> 
  select(c(1, 43:46, 277:280)) |> 
  distinct() |> 
  arrange(participantID)

#present an overview of the tibble

knitr::kable(head(mar10_commu), format = "pipe") 
```

##### **Monthly survey April, 2010**

Case for this data set is (participant, month).

*Big 5*

Big 5 included all classic 44 questions of the Big 5 personality model. This scale was recorded only once.

```{r}
apr_big5 = month1004_df |> 
  select(c(1, 11:54)) |> 
  distinct() |> 
  arrange(participantID)

#present an overview of the tibble

knitr::kable(head(apr_big5), format = "pipe") 
```

*Group participation*

```{r}
apr_group = month1004_df |> 
  select(c(1, 55:114)) |> 
  distinct() |> 
  arrange(participantID)

#present an overview of the tibble

knitr::kable(head(apr_group), format = "pipe") 
```

##### **Monthly survey May, 2010**

Case for this data set is (participant, month).

*Group participation*

```{r}
may_group = month1005_df |>
  select(c(1, 73:132)) |> 
  distinct() |> 
  arrange(participantID)

#present an overview of the tibble

knitr::kable(head(may_group), format = "pipe") 
```

*Well-being*

Well-being measures participants' monthly psychological state by agreeing or disagreeing sets of statements. This scale was recorded only once.

```{r}
may_well_being = month1005_df |> 
  select(c(1, 32:72)) |> 
  distinct() |> 
  arrange(participantID)

#present an overview of the tibble

knitr::kable(head(may_well_being), format = "pipe") 
```

##### **Monthly survey October, 2010**

Case for this data set is (participant, month).

*Aerobic exercise actual*

Aerobic exercise actual records the exact hours of participants' aerobic activities. This variable was recorded only once.

```{r}
oct_exe_act = month1010_df |>    
  select(c(1, 62)) |>    
  distinct() |>    
  arrange(participantID)

#present an overview of the tibble

knitr::kable(head(oct_exe_act), format = "pipe") 
```

*Sleep actual*

Sleep actual records the exact hours of participants' sleeping duration. This variable was recorded only once.

```{r}
oct_sleep_act = month1010_df |> 
  select(c(1, 80)) |> 
  distinct() |> 
  arrange(participantID)

#present an overview of the tibble

knitr::kable(head(oct_sleep_act), format = "pipe") 
```

##### **Monthly survey March, 2011**

Case for this data set is (participant, month).

*Communication with friends*

```{r}
mar11_commu = month1103_df |> 
  select(c(1, 26:30)) |> 
  distinct() |> 
  arrange(participantID)

#present an overview of the tibble

knitr::kable(head(mar11_commu), format = "pipe") 
```

### 2.1.2 Location

Location every 30 minutes (`Location.csv`), with each record including the participant ID associated with the data-collecting mobile phone (`participantID`), the date-time when the location is recorded (`date`), the the 68% confidence offset of the estimation from true location(`accuracy`, in meters), and the estimated location affine-transformed (`x`, `y`).

```{r}
#create a tibble for Location.csv, and have a preview of the data set

location_df = read.csv("Location.csv") |> arrange(participantID)

knitr::kable(head(location_df), format = "pipe") 
```

We tried to define the primary key, however, even if we combine all the column names together, the combination itself still cannot identify every single row uniquely, as it is shown in the following code:

```{r}
#check whether (participantID, date, accuracy, x, y) is the primary key

location_df |> 
  count(participantID, date, accuracy, x, y) |> 
  filter(n>1) |> 
  head()
```

We deemed that the duplicates were produced due to faulty data collection methods after reviewing the data description. Thus, we decided to remove duplicated data that has the exact value under every variable:

```{r}
location_df = location_df |> distinct()
```

In the new data frame, we found that (`participantID`, `date`) is the primary key, as it is shown in the following code:

```{r}
#check whether (participantID, date) is the primary key

location_df |> 
  count(participantID, date) |> 
  filter(n>1)
```

In the final data frame, for the case participant ID, no characteristic was involved. For the case (`participantID`, `date`), we created a separate tibble according to principle 1:

```{r}
date_coordinates = location_df |> 
  select(participantID, date, x, y) |> 
  distinct() |> 
  arrange(participantID)

#present an overview of the tibble

knitr::kable(head(date_coordinates), format = "pipe") 
```

## 2.2 Principle 2: Each observation must have its own row.

For principle 2, the researcher needed to ensure that no multiple observations were stacked into the same row. In order to achieve that, the researcher first examined all 11 tibbles created from the first step to spot repeated information. Then, the researcher deleted repeated information among different rows.

### 2.2.1 Monthly Survey

#### **Initial Survey/March, 2010**

*Demographics*

```{r}
# check to count the distinct values within Gender

knitr::kable(count(mar_demo, Gender), format = "pipe") 
```

While only two genders (female and male) were represented in the mar_demo participants, the data frame included multiple different expressions of the same terms. Thus, the researcher decided to change the values for the Gender column to become consistent.

```{r}
# recode the values in Gender to be consistent and use the same expressions

mar_demo <- mar_demo |> 
  transmute(participantID = participantID,
            Gender =
              recode(Gender,
                     "F" = "Female",
                     "f" = "Female",
                     "female" = "Female",
                     "M" = "Male",
                     "m" = "Male",
                     "male" = "Male")) |> 
  distinct()

# check to count the distinct values within Gender after cleaning

knitr::kable(count(mar_demo, Gender), format = "pipe") 
```

From the participant ID, the researcher knew that there were in total 86 + 56 = 142 participants, but only 70 +71 = 141 rows were found in mar_demo. The researcher then explored the missing participant.

```{r}
# check for the missing participant from the demographics df.

all_participants <- c(paste0("fa10-01-", sprintf("%02d", 1:86)),
                      paste0("sp10-01-", sprintf("%02d", 1:56)))
setdiff(all_participants, mar_demo$participantID)
```

The missing participant "fa10-01-84" was present in other survey logs. Only their demographic info, specifically gender, was missing. In that case, in order to preserve the correct sample size in this demographic data frame, the researcher decided to add a row with the participant ID and NA value for the Gender column.

```{r}
# add the row for the missing participant fa10-01-84.

mar_demo <- mar_demo |> 
  add_row(Gender = NA, participantID = 'fa10-01-84') |> 
  arrange(participantID)
```

*Group Participation*

```{r}
# check the head of the data frame
knitr::kable(head(mar_group), format = "pipe") 
```

After preliminary examination of the data frame mar_group, the researcher discovered that there were many rows with NA values for all columns (except for the participant ID). These rows essentially provide no information and were thus deleted.

```{r}
# delete rows with all NA values because they provide essentially no information.

mar_group <- mar_group |> 
  filter(!if_all(-participantID, ~is.na(.)))
```

There should be 56 participants with answers for Group Participation, but only 55 rows of data were there. The researcher then explored the missing participant.

```{r}
# explore which participant is missing from the data frame.

all_participants <- c(paste0("sp10-01-", sprintf("%02d", 1:56)))
setdiff(all_participants, mar_group$participantID)
```

In addition, the researcher noticed that the NA values were not consistently represented in all cells: some cells were left blank, which essentially means the same thing as NA. So, the researcher decided to recode the blank cells to retain consistency.

```{r}
# fill all empty cells with NA values

mar_group <- mar_group |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  distinct()
```

Notably, upon further inspect of the original tibble, the researcher discovered that column headers contain multiple types of information: besides the belonging/involvement variable, they also contain group number (1-6) and group type (formal/informal). This is in violation with principle 2 which stated "each observation must have its own row."

First, the researcher investigated the belonging columns. The `pivot_longer`function was adopted to transition from the wide format to a longer one, with new column headers `Formal_Or_Informal` and `Group_Number` storing the names either preceding or appended to the `Group_Belonging`variable. The problem lays with the fact that the variable `Group_Number` does not store the value of the *i*th group for each of the *n* participants, but the value of the *i*th formal or *i*th informal group. This is lacking in clarity. We thus simply count the rows which share the same `participantID`.

```{r}
mar_group_belonging <- mar_group |>
  select(participantID, contains("belong.to"))|>
  pivot_longer(
    cols = 'Formal.groups.belong.to..Group.1.': 'Informal.groups.belong.to..Group.6.',
    names_to = c("Formal_or_Informal", "Group_Number"),
    names_pattern = "(Formal|Informal)\\.groups\\.belong\\.to\\.\\.Group\\.(\\d+)\\.",
    values_to = "Group_Belonging",
    values_drop_na = TRUE) |>
  group_by(participantID) |>
  mutate(
    Group_Number = row_number()
  ) |>
  ungroup() 
```

Second, the same procedure was applied to the involvement columns.

```{r}
mar_group_involvement <- mar_group |>
  select(participantID, contains(".involvement"))|>
  pivot_longer(
    cols = 'Formal.group.involvement..Group.1.': 'Informal.group.involvement..Group.6.',
    names_to = c("Formal_or_Informal", "Group_Number"),
    names_pattern = "(Formal|Informal)\\.group\\.involvement\\.\\.Group\\.(\\d+)\\.",
    values_to = "Group_Involvement",
    values_drop_na = TRUE) |>
    group_by(participantID) |>
    mutate(
      Group_Number = row_number()
    ) |>
    ungroup() 
```

Lastly, we left joined the two artificially split tibbles to recreate the full one. `mar_group_involvement` features `r nrow(mar_group_belonging)` observations, and `mar_group_belonging` featured `r nrow(mar_group_involvement)`, meaning that certain belonged-to groups had missing values for involvement. By left-joining on belonging, we captured those cases as well.

```{r}
mar_group <- left_join(
  mar_group_belonging,         
  mar_group_involvement,     
  by = c("participantID", "Group_Number", "Formal_or_Informal")
) |>
  select(participantID, Group_Number, Group_Belonging, Group_Involvement, Formal_or_Informal)
```

The clean tibble then looked like:

```{r}
knitr::kable(head(mar_group,5), format = "pipe") 
```

*Communication with friends*

Glimpse the communication table before principle two is applied. Frequencies are measured in terms of participant *and* channel. So that each such observation can have its own row, this tibble too must be pivoted longer, in accordance with principle two.

```{r}
mar10_commu[mar10_commu == ""] <- NA
knitr::kable(head(mar10_commu,5), format = "pipe") 
```
In addition, the researcher discovered that participants with ID starting "sp" and "fa" were asked two different sets of questions with overlapping questions. The participants sp10-01-01 to sp10-01-56 answered questions on channels: phone, text, email, and chat, while participants fa10-01-01 to fa10-01-86 answered questions on channels: phone, Skype, text, and chat.

Thus, the researcher decided to combine the overlapping columns together. 

```{r}
# combine the three columns that essentially asked the same questions but on different waves of participants

mar10_commu <- mar10_commu |> 
  transmute(participantID = participantID,
            Communicate.remote.friends..email = Communicate.remote.friends..email,
            Communicate.remote.friends..internet.phone..ex..Skype. = Communicate.remote.friends..internet.phone..ex..Skype.,
            Communicate.remote.friends..phone..landline.or.cell.phone. = 
           coalesce(Communicate.remote.friends..phone, Communicate.remote.friends..phone..landline.or.cell.phone.),
         Communicate.remote.friends..text.message.SMS =
           coalesce(Communicate.remote.friends..text.message.SMS, Communicate.remote.friends..text),
         Communicate.remote.friends..online.chat.instant.message = 
           coalesce(Communicate.remote.friends..online.chat.instant.message, Communicate.remote.friends..chat),
         )
```

We then simply pivot from wide format to long, filling in the values with the strings succeeding 'friends' and replacing the dots with blank spaces for clarity.

```{r}
mar10_commu <- mar10_commu |>
  pivot_longer(
    cols = -participantID,
    names_to = "Channel",
    names_pattern ="Communicate\\.remote\\.friends\\.\\.(.*)",
    values_to = "Frequency") |>
  mutate(
    Channel = str_replace_all(Channel, "\\.", " ") 
  )
```

To make the NA values be represented consistently across cells, the researcher coded to fill empty cells with NA values.

```{r}
# fill all empty cells with NA values

mar10_commu <- mar10_commu |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  distinct()
```

Now the output looks like this:

```{r}
knitr::kable(head(mar10_commu,5), format = "pipe") 
```

#### **Monthly survey April, 2010**

*Big 5*

```{r}
# fill all empty cells with NA values

apr_big5 <- apr_big5 |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  distinct()
```

```{r}
# explore how many participants are missing from the apr_big5 df.

all_participants <- c(
  paste0("sp10-01-", sprintf("%02d", 1:56))
)
setdiff(all_participants, apr_big5$participantID) |> 
  length()
```

*Group participation*

```{r}
# fill all empty cells with NA values

apr_group <- apr_group |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  distinct()

# delete rows with all NA values because they provide essentially no information.

apr_group <- apr_group |> 
  filter(!if_all(-participantID, ~is.na(.)))
```

```{r}
# explore how many participants are missing from the apr_group df.

all_participants <- c(
  paste0("sp10-01-", sprintf("%02d", 1:56))
)
setdiff(all_participants, apr_group$participantID) |> 
  length()
```

```{r}
# explore how many participants are missing from the apr_group df.

all_participants <- c(
  paste0("sp10-01-", sprintf("%02d", 1:56))
)
setdiff(all_participants, apr_group$participantID) |> 
  length()
```

The researcher spotted several repeated values in the columns and decided to pivot the columns into multiple rows for each different observation.

```{r}
# change the data type into character

apr_group <- apr_group |>
  mutate(across(-participantID, as.character))
```
Group participation for the months following the initial survey is slightly more complex: we now have groups participants have stopped being a part of, groups participants have joined (with a new involvement variable) and groups participants had previously been a part of, but for which the level of involvement has changed. Still, a similar type of problematic data structure persists, as with the previous group participation tibble.

```{r}
knitr::kable(head(apr_group,5), format = "pipe")
```

First, we handle all columns not partaking to involvement. The first part of the column marks the type of group, the middle the action the participant has engaged in (joined/stopped/alerted participation), and the last the group number, which we must again change to number of rows associated with one `participantID`. We drop the *NA*s so we obtain only the groups for which an action and description was ascribed.

```{r}
apr_group_descript <- apr_group|>
  select(participantID, !contains("involvement"))|>
  pivot_longer(
    cols = -participantID,
    names_to = c("Formal_or_Informal", "Group_Action", "Group_Number"),
    names_pattern = "(Formal|Informal)groups(.*)\\.Group(\\d+)",
    values_to = "Group_Belonging",
    values_drop_na = TRUE)|>
  group_by(participantID)|>
  mutate(
    Group_Number = row_number())
```

We handle involvement in a similar fashion. We drop *NA*s so we obtain only the groups for which a level of involvement was specified.

```{r}
apr_group_involvement <- apr_group|>
  select(participantID, contains("involvement"))|>
  pivot_longer(
    cols = -participantID,
    names_to = c("Formal_or_Informal", "Involvement_Type", "Group_Number"),
    names_pattern = "(Formal|Informal)groups(.*)\\.Group(\\d+)",
    values_to = "Group_Involvement",
    values_drop_na =TRUE)|>
  group_by(participantID)|>
  mutate(
    Group_Number = row_number())
```

Finally, we left-join the two on `apr_group_descript`, so we do not miss out the `r nrow(apr_group_descript) - nrow(apr_group_involvement)` values for which a group was described, but an involvement level not specified.

```{r}
apr_group <- left_join(
  apr_group_descript,         
  apr_group_involvement,     
  by = c("participantID", "Group_Number", "Formal_or_Informal")
) |>
  select(participantID, Group_Number, Group_Action, Group_Belonging, Group_Involvement, Formal_or_Informal)


```

The final tibble now looks like:

```{r}
knitr::kable(head(apr_group,5), format = "pipe") 
```

#### **Monthly survey May, 2010**

*Group participation*

```{r}
# fill all empty cells with NA values

may_group <- may_group |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  distinct()

# delete rows with all NA values because they provide essentially no information.

may_group <- may_group |> 
  filter(!if_all(-participantID, ~is.na(.)))

# explore how many participants are missing from the apr_group df.

all_participants <- c(
  paste0("sp10-01-", sprintf("%02d", 1:56))
)
setdiff(all_participants, may_group$participantID) |> 
  length()
```
The same procedure as above was applied to this tibble, as the data structure was identical to April, 2010. 

```{r}
#Pivoting group description and action tibble
may_group_descript <- may_group|>
  select(participantID, !contains("involvement"))|>
  pivot_longer(
    cols = -participantID,
    names_to = c("Formal_or_Informal", "Group_Action", "Group_Number"),
    names_pattern = "(Formal|Informal)groups(.*)\\.Group(\\d+)",
    values_to = "Group_Belonging",
    values_drop_na = TRUE)|>
  group_by(participantID)|>
  mutate(
    Group_Number = row_number())
#Pivoting involvement tibble
may_group_involvement <- may_group|>
  select(participantID, contains("involvement"))|>
  pivot_longer(
    cols = -participantID,
    names_to = c("Formal_or_Informal", "Involvement_Type", "Group_Number"),
    names_pattern = "(Formal|Informal)groups(.*)\\.Group(\\d+)",
    values_to = "Group_Involvement",
    values_drop_na =TRUE)|>
  group_by(participantID)|>
  mutate(
    Group_Number = row_number())
#Putting the two separated tibbles together
may_group <- left_join(
  may_group_descript,         
  may_group_involvement,     
  by = c("participantID", "Group_Number", "Formal_or_Informal")
) |>
  select(participantID, Group_Number, Group_Action, Group_Belonging, Group_Involvement, Formal_or_Informal)

```

*Well-being*

```{r}
# fill all empty cells with NA values

may_well_being <- may_well_being |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  distinct()

# delete rows with all NA values because they provide essentially no information.

may_well_being <- may_well_being |> 
  filter(!if_all(-participantID, ~is.na(.)))

# explore how many participants are missing from the may_well_being df.

all_participants <- c(
  paste0("sp10-01-", sprintf("%02d", 1:56))
)
setdiff(all_participants, may_well_being$participantID) |> 
  length()
```

#### **Monthly survey October, 2010**

*Aerobic exercise actual*

```{r}
# fill all empty cells with NA values

oct_exe_act <- oct_exe_act |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  distinct()

# delete rows with all NA values because they provide essentially no information.

oct_exe_act <- oct_exe_act |> 
  filter(!if_all(-participantID, ~is.na(.)))

# explore how many participants are missing from the oct_exe_act df.

all_participants <- c(
  paste0("sp10-01-", sprintf("%02d", 1:56)),
  paste0("fa10-01-", sprintf("%02d", 1:40))
)
setdiff(all_participants, oct_exe_act$participantID) |> 
  length()
```

*Sleep actual*

```{r}
# fill all empty cells with NA values

oct_sleep_act <- oct_sleep_act |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  distinct()

# delete rows with all NA values because they provide essentially no information.

oct_sleep_act <- oct_sleep_act |> 
  filter(!if_all(-participantID, ~is.na(.)))

# explore how many participants are missing from the oct_sleep_act df.

all_participants <- c(
  paste0("sp10-01-", sprintf("%02d", 1:56)),
  paste0("fa10-01-", sprintf("%02d", 1:40))
)
setdiff(all_participants, oct_sleep_act$participantID) |> 
  length()
```

#### **Monthly survey March, 2011**

*Communication with friends*

```{r}

# delete rows with all NA values because they provide essentially no information.

mar11_commu <- mar11_commu |> 
  filter(!if_all(-participantID, ~is.na(.)))

# fill all empty cells with NA values

mar11_commu <- mar11_commu |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  distinct()

# explore how many participants are missing from the mar11_commu df.

all_participants <- c(
  paste0("sp10-01-", sprintf("%02d", 1:56)),
  paste0("fa10-01-", sprintf("%02d", 1:40))
)
setdiff(all_participants, mar11_commu$participantID) |> 
  length()
```
The initial tibble is very similar to the 2010 communication one, except that the pattern the column names follow is slightly different.

```{r}
knitr::kable(head(mar11_commu,5), format = "pipe") 

```

We thus update the `name_pattern` argument. 

```{r}
mar11_commu <- mar11_commu |>
  pivot_longer(
    cols = -participantID,
    names_to = "Channel",
    names_pattern ="Communicateremotefriends\\.(.*)",
    values_to = "Frequency") |>
  mutate(
    Channel = str_replace_all(Channel, "\\.", " ") 
  )
```

Result:

```{r}
knitr::kable(head(mar11_commu,5), format = "pipe") 

```
### 2.1.2 Location

```{r}
# delete possible repeated rows

date_coordinates <- date_coordinates |> 
  distinct(participantID, date, x, y, .keep_all = TRUE)

# fill all empty cells with NA values

date_coordinates <- date_coordinates |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  distinct()

# delete rows with all NA values because they provide essentially no information.

date_coordinates <- date_coordinates |> 
  filter(!if_all(-participantID, ~is.na(.)))

# explore how many participants are missing from the date_coordinates df.

all_participants <- c(
  paste0("sp10-01-", sprintf("%02d", 1:56)),
  paste0("fa10-01-", sprintf("%02d", 1:40))
)
setdiff(all_participants, date_coordinates$participantID) |> 
  length()
```

## 2.3 Principle 3: Each variable must have its own column. 

According to the third principle, each variable must have its own column, and each column must be a variable. This implies, also, that each variable must be stored in its unique, separate column. Headers must be variable names, and not data; variables must not appear in both rows and columns.

After inspecting each of our designated tibbles, it became apparent that none were in violation of principle three. As none of our tibbles were of extensive size, using `pivot_longer` did not produce further issues with the data structure, such as spreading a variable across several columns. In general, the heftiest issues with the semantics of this dataset had to do with principle two violations, such as in the case of `mar_group` and `apr_group`. Thus, no situations were the use of `pivot_wider` was appropriate appeared.

### 2.4.2 Location

In the data set of location (`date_coordinates`*)*, the date and time are combined into a single cell. Since participants' locations were recorded at multiple time points within the same day, keeping date and time in one cell makes it difficult to perform analyses based solely on "*day"* or "*time of day"*. Therefore, we need to separate the datetime column into two distinct columns, **Date** and **Time**. This step is necessary because the date and the time represent two distinct types of information, and placing them in separate cells improves the clarity of the data set.

We use `separate()` function to separate these two values.

```{r}
#separate the datetime column into two colunms 
date_coordinates <- date_coordinates |> 
  separate(date, c('Date', 'Time'), " ")

#present an overview of the tidied tibble

knitr::kable(head(date_coordinates), format = "pipe")
```

# 3. Description of Tidy Data

The tidy data sets includes participant id as well as all 7 variables relevant to our research questions, namely demographics, group participation, communicating with friends, big 5, well-being, aerobic exercise actual, and sleep actual. Survey data of all variables mentioned was collected during March to may, 2010, as well as October, 2010 and March 2011. Regarding participants, 56 participants joined the study from March, 2010 (Spring participants, participantID starting with sp-), and 86 participants joined the study from October,2010 (Fall participants, participantID starting with fa-). Therefore, variables that were recorded only once between March and May, 2010 only contains survey responses from spring participants.

## 3.1 Doubt regarding data quality

While we were trying to clean the data sets, several doubts were raised.

First of all, issues in data collecting. Regarding self-reporting items, researchers did not restrict numbers of submissions per participant based on the IP address. As a result, participants might open the same survey for twice with all of the trials being recorded, and one of them consisting of NA values entirely. Regarding auto-recorded items, such as location, data points with the same participant id, the same recording date and time, and the same coordinates were sometimes also recorded twice. 

Second, survey items that measures the same variable changed across different waves of the study. For example, for the variable communicating with friends, channels were named different according to the SurveyMonthlyQuestionKey.csv:

In March, 2010, for spring participants, channels names were: phone, text, email and chat. For fall participants, however, channels names were: land-line/cellphone, Skype, text message/SMS, and online chat/instant message.

In March, 2011, number of items changed from 4 to 5, where they combined all the items measured before, namely: land-line/cellphone, Skype, text message/SMS, online chat/instant message and email. They completely discarded the items that were first outlined in the initial survey, instead they just added email as an extra item in the follow up survey in March, 2011. This mismatch of survey items could possibly decrease the internal validity of the study in this case.




















